Parallel Simulation of Adaptive Random Boolean Networks

### 2016-07-15 Paper/paper.tex
Final draft

\def\procediaConference{99th Conference on Topics of
	Superb Significance (COOL 2014)}

@@

	As RBNs are applied to model biological structures, it is interesting to study the evolution of RBNs. Core of the evolution process is a constant searching for \footnote{Tu sie nie zgadzam, ze powinno być of. Istnieja 2 zwroty: 'search for' oraz 'in search of'. Ten drugi znaczy 'w poszukiwaniu'.} the most optimal configuration \cite{aldana2003boolean,sole1996extinction,zimmermann2004coevolution}. 
	An RBN modification that incorporates system evolution is {\it adaptive} RBN (ARBN) \cite{mlb,soc,haruna14relationship,Gorski2016}. 
	In RBNs nodes can be divided into two groups of active and frozen nodes. Nodes of the former group change their state during network attractor. On the other hand, the frozen nodes keep their state. 
	The basis of ARBN is a natural idea that overactive nodes should be quietened down and underactive (frozen) ones should be stimulated. This concept is included in an activity-dependent rewiring rule (ADRR) \cite{bornholdt2000topological}, which is as follows: 
	if the node’s mean state $\langle\sigma_i\rangle$ during the network attractor is $0$ or $1$, then this node
	is considered to be frozen and one new incoming edge is
	added to this node. Otherwise, this node is considered to be active and one of
	its incoming edges is randomly chosen and deleted.

@@

	In this paper, we perform ARBN simulations for larger networks with up to $1~000$ nodes. 
	In such structures the difficulty is in finding the attractor because transient and attractor lengths grow quickly with network size. 
	In previous work there have been only few results with such network sizes. In original paper \cite{mlb} the full model analysis is for the network of size up to $400$ nodes\footnote{TODO: Later it might be emphasized that LiuBassler results must have skipped most (significant number) of the attractors.}. 
	In our previous work, where we analyze modular Boolean networks with separate ARBNs connected by a number of interlinks \cite{Gorski2016}, the studied maximal network size is $80$ nodes. 
	Here, we do not include modularity into ARBN model. 
	This research is a step towards large scale simulations of modular ARBNs \cite{Gorski2016}. To investigate large ARBNs, we design a parallel algorithm for simulation of ARBN's evolution and combine it with several algorithmic improvements.

	The remaining part of the paper is organized as follows. Section 2 gives an overview of a related work on algorithmics of RBN modeling. Section 3 presents a description of algorithms and techniques which were used to speed up simulations (including an algorithm for parallel simulation of adaptive RBNs). Section 4 (an experimental investigation) has two-fold purpose: (i) to study the performance of the parallel algorithm and the influence of proposed modifications on simulation results, and (ii) to study the dynamics of evolution of non-modular ARBNs up to 1000 nodes.

	%------------------------------------------------------------------------------
	\section{Related Work}
	\label{sect:related-work}
	In recent years, a variety of RBNs has been extensively studied by the means of computer simulation. Basically, an initial topology, a set of boolean functions and an updating scheme (synchronous or asynchronous) \cite{gershenson04} are fixed. In contrast, adaptive RBN can evolve by changing links and/or boolean functions of nodes. Bornholdt and Sneppen in \cite{bornholdt1998neutral} introduced an algorithm for RBN's evolution under the constraint of continuity. In the algorithm, a degree of a randomly chosen node is modified only when a network with this modification reaches the same attractor as the mother system did. Paczuski, Bassler and Corral in \cite{paczuski2000self} consider nodes of RBN as competing agents in a market. The evolution of network is maintained by a random change of the poorest player's strategy at each epoch of simulation. ADDR as a mechanism for transition of RBN to a critical state was proposed in \cite{bornholdt2000topological} and then extended in \cite{mlb} for annealed and quenched versions of a model. Both in  \cite{bornholdt2000topological} and \cite{mlb} ADDR changes a single randomly chosen node per iteration according to its average state on the attractor; the difference is that in \cite{mlb} boolean functions of nodes can also be changed at the end of the epoch.

	Simulation of a single epoch for adaptive RBN consists of two steps, search for an attractor and modification of a network according to chosen rules. Depending on the goal of simulation, it could be necessary to find a single attractor for a given initial state or all the attractors of a system. Zhang et. al. in \cite{zhang2007algorithms} propose algorithms to find singleton attractors of RBN using gene ordering. To find a single attractor of an arbitrary length, there are two main algorithms: Knuth's algorithm \cite{knuth1981art} (the example of its usage for RBN's simulation can be found in \cite{bhattacharjya1996median}) and the algorithm described in \cite{mlb} (we denote it as Liu-Bassler). We include the brief description of these algorithms in Section 3 of this paper as they serve as a basis for our parallel algorithm. The second category includes algorithms to find all attractors for a given network. These algorithms usually reformulate basic task as binary decision diagrams, satisfiability (SAT) or aggregation problems. For the asynchronous update mode, there was proposed an approach based on an exhaustive enumeration of subspaces of a network space implemented in Inet software package \cite{berntenis2013detection}. Some of the algorithms of this type allow parallel implementation (e.g. \cite{guo2014parallel}). Parallelization is performed by calculating different attractors (or sets of attractors) for the same network on different processors, and is not appropriate for ADDR having data dependency between epochs.

@@

	Due to the exponential slowing down by long attractor periods for initial connectivity $K_{ini} \geq 2$, usually it is necessary to put the restrictions on a maximum number of iterations to be examined during a single search for attractor (it is denoted as $T_{max}$). In \cite{bornholdt2000topological}, authors use $T_{max} = 1000$ which allows them to study networks up to 1024 nodes. In \cite{mlb}, $T_{max}$ is set up to 100 000, and the maximum size of a studied network is equal to 400 nodes. Increase of a size of a network with fixed $T_{max}$ leads to skipping a significant number of the attractors. However, authors of \cite{bornholdt2000topological} and \cite{mlb} do not consider a percentage of attractors found, and how the results of simulation are influenced by skipped attractors.

@@

	Modeling the evolution of ARBN larger than several dozens of nodes is hampered by a significant growth of transient and attractor lengths. This growth extremely prolongs the time of a repeated calculation of nodes' states during search for an attractor (which is the most time consuming part of a modeling). To speed up modeling for large ARBNs, we propose to combine algorithms and techniques described in this Section. We use existing sequential algorithms for attractor's search with a limitation on a maximum number of iterations as a basis of the parallel algorithm, and extend ADDR to support larger number of nodes to be rewired.

@@

The modified algorithm of ARBN evolution consists of a set of consecutive epochs. In each epoch the network attractor is found and $N_{rew}$ nodes are rewired according to the modified ADRR. The maximum length of attractors $\mu_{max}$ is limited ether by predefined checkpoints in Liu-Bassler algorithm, or artificially in Knuth's algorithm in order to speed up the computations as described in Section \ref{sect:algorithm}. Due to this limitation only a fraction $R$ of the attractors is found. During the evolution process the system reaches a dynamical equilibrium, which can be observed by measuring network's mean in-degree connectivity $K$, defined in Eq. \ref{eq:connectivity}. As the network evolves, $K$ tends to some steady-state value $K_{ss}$.

@@

	Intuitively, Liu-Bassler algorithm of attractor's search would have shown better performance than the Knuth's one, as the latter requires at least $3 \mu$ state updates to find an attractor of length $\mu$. But in order to use Liu-Bassler algorithm one needs to choose the checkpoints correctly, which requires an a-priori knowledge of attractor and transient period distributions. This distribution depends on a network size, connectivity and structure (e. g. modular networks tend to have shorter attractors than non-modular ones \cite{Gorski2016}). We obtained an empirical distribution of attractor and transient lengths for different kinds of networks in our preliminary experiments with Knuth's algorithm. The checkpoints $T=[100, 200, 1000, 2000, ... \mu_{max}, 2 \mu_{max}]$ seem to be the best trade-off between the number of state updates and the percentage of attractors found. However, with these checkpoints the difference in the number of state updates between the two algorithms is negligible. We used Knuth's algorithm in our future experiments, as it is more general and requires no a-priori knowledge of the network's behaviour.

	Both algorithms were implemented using C++ programming language. The implementation is cross-platform and requires a GPU with CUDA compute capability at least 1.2. The representation of RBNs' state update function defined in Eq. \ref{eq:spmv}, \ref{eq:gather} allows for experimenting with different SpMV algorithms. These algorithms mainly aim at maximizing memory throughput and data locality, thus reducing the overall GPU kernel execution time. We chose compressed sparse-row (CSR) matrix storage format and implemented an sparse matrix-vector multiplication kernel as described in \cite{bell2008efficient}, as it shows the best performance on unstructured matrices.

@@

	\paragraph{Speedup obtained by GPU acceleration.} We measured the execution times of both sequential and parallel versions of state updates. We used Liu-Bassler algorithm with checkpoints $T$ = $\{100,$ $200,$ $1000,$ $2000,$ $...,$ $10^7,$ $2 \cdot 10^7\}$, $\mu_{max} = 10^7$. To decrease the number of random samples required to compare the versions, we did not stop the epoch as soon as an attractor was found. Instead, we performed all $2 \cdot 10^7$ state updates per epoch. Such simulation corresponds to the worst-case scenario when no attractors were found, i. e. $R = 0\%$. The results presented in Figure \ref{fig:speedup} indicate that the GPU implementation is more efficient than the serial one for networks consisting of more than $200$ nodes.

	\begin{figure}[ht!]
		\begin{minipage}[t]{0.45\textwidth}
			\includegraphics[width=1.0\textwidth]{plots/speedup}
			\caption{Execution time of the GPU algorithm compared to the sequential one}
			\label{fig:speedup}
		\end{minipage}\hfill
		\begin{minipage}[t]{0.45\textwidth}
			\includegraphics[width=1.0\textwidth]{plots/n_rew}
			\caption{Evolution of mean connectivity for different $N_{rew}$; $N=1000$}
			\label{fig:nrew}
		\end{minipage}
	\end{figure}

	\paragraph{Rewiring more than one node per epoch.} In order to determine how $N_{rew}$ influences the rate of convergence to a dynamical equilibrium, we performed $1000$ epochs of evolution with $N=1000$ and $N_{rew}=[1, 5]$. The results of these simulations are shown in Figure \ref{fig:nrew}. With increased value of $N_{rew}$ mean connectivity grows faster and the system reaches steady state at approximately 500th epoch for $N_{rew}=5$. Therefore, we conclude that by rewiring $N_{rew} > 1$ nodes per epoch one can lower the number of epochs required to reach steady state. We limited the attractor length to $\mu_{max}=10^7$ for these experiments.

	\paragraph{Limiting the maximum attractor length.} Introducing a limit on attractor length leads to a situation when an attractor can not be found. In these cases $N_{rew}$ nodes are rewired according to their activity during the last $\mu_{max}$ state updates. These states can belong to either a transient period or an attractor. However, the results of our experiments, presented in Tables \ref{table:percentage_of_attractors}, \ref{table:kss}, indicate that low percentage of attractors found does not influence the steady state connectivity obtained as a result of the evolution process.

@@

	Using the approaches discussed above we were able perform simulations on networks consisting of up to 1000 nodes. Plots of the evolution of mean in-degree connectivity and the empirical distribution of attractor lengths are shown in Figures \ref{fig:evolution} and \ref{fig:attractors} respectively. The peaks near $\mu=10^7$ in Figure \ref{fig:attractors} are caused by the limit on the attractor length $\mu_{max} = 10^7$.

	\begin{figure}[h]
		\begin{minipage}[t]{0.45\textwidth}
			\includegraphics[width=1.0\textwidth]{plots/evolution}
			\caption{Evolution of mean network connectivity}
			\label{fig:evolution}
		\end{minipage}\hfill
		\begin{minipage}[t]{0.45\textwidth}
			\centering
			\includegraphics[width=1.0\textwidth]{plots/attractors}
			\caption{Empirical distribution of attractor lengths}
			\label{fig:attractors}
		\end{minipage}
	\end{figure}
	%------------------------------------------------------------------------------
	\section{Conclusion}
	The enormous size of RBN ensemble makes them appropriate to model various complex systems but also raises significant challenges for their numerical investigation. Even a problem of search for a single attractor of RBN is computationally hard for a networks larger than several dozens of nodes because of an exponential growth of a number of iterations with an increase of a network size.

@@

	Using the algorithm, we performed simulation of networks up to 1000 nodes on a hybrid cluster with NVidia graphic cards. The parallel algorithm demonstrates increasing performance with a growth of a network size (for 1000-nodes network it is approximately 8 times faster than a sequential one). The results show that proposed modification of the rewiring rule allows to reduce a number of epochs to reach a steady state making calculations less time-consumable. We also present the results for a percentage of attractors found for different combinations of sizes of networks and maximum number of iterations examined during search for a single attractor. These values together with the results on mean steady state connectivity can be used for a choice of values of checkpoints in the sequential algorithms of attractor's search.

>>>>

@@

	As RBNs are applied to model biological structures, it is interesting to study the evolution of RBNs. Core of the evolution process is a constant searching for %\footnote{Tu sie nie zgadzam, ze powinno być of. Istnieja 2 zwroty: 'search for' oraz 'in search of'. Ten drugi znaczy 'w poszukiwaniu'.}
	the most optimal configuration \cite{aldana2003boolean,sole1996extinction,zimmermann2004coevolution}. 
	An RBN modification that incorporates system evolution is {\it adaptive} RBN (ARBN) \cite{mlb,soc,haruna14relationship,Gorski2016}. 
	In RBNs nodes can be divided into two groups of active and frozen nodes. Nodes of the former group change their state during network attractor. On the other hand, the frozen nodes keep their state. 
	The basis of ARBN is a natural idea that overactive nodes should be quietened down and underactive (frozen) ones should be stimulated. This concept is included in an activity-dependent rewiring rule (ADRR) \cite{bornholdt2000topological}, which is as follows: 
	if the node’s mean state $\langle\sigma_i\rangle$ during the network attractor is $0$ or $1$, then this node
	is considered to be frozen and one new incoming edge is
	added to this node. Otherwise, this node is considered to be active and one of
	its incoming edges is randomly chosen and deleted.

@@

	In this paper, we perform ARBN simulations for larger networks with up to $1~000$ nodes. 
	In such structures the difficulty is in finding the attractor because transient and attractor lengths grow quickly with network size. 
	In previous work there have been only few results with such network sizes. In original paper \cite{mlb} the full model analysis is for the network of size up to $400$ nodes
	%\footnote{TODO: Later it might be emphasized that LiuBassler results must have skipped most (significant number) of the attractors.}. 
	In our previous work, where we analyze modular Boolean networks with separate ARBNs connected by a number of interlinks \cite{Gorski2016}, the studied maximal network size is $80$ nodes. 
	Here, we do not include modularity into ARBN model. 
	This research is a step towards large scale simulations of modular ARBNs \cite{Gorski2016}. To investigate large ARBNs, we design a parallel algorithm for simulation of ARBN's evolution and combine it with several algorithmic improvements.

	The remaining part of the paper is organized as follows. Section 2 gives an overview of a related work on algorithmics of RBN modeling. Section 3 presents a description of algorithms and techniques which were used to speed up simulations (including an algorithm for parallel simulation of adaptive RBNs). Section 4 (an experimental investigation) has two-fold purpose: (i) to study the performance of the parallel algorithm and the influence of proposed modifications on simulation results, and (ii) to study the dynamics of evolution of non-modular ARBNs up to $1~000$ nodes.

	%------------------------------------------------------------------------------
	\section{Related Work}
	\label{sect:related-work}
	In recent years, a variety of RBNs has been extensively studied by the means of computer simulation. Basically, an initial topology, a set of boolean functions and an updating scheme (synchronous or asynchronous) \cite{gershenson04} are fixed. In contrast, adaptive RBN can evolve by changing links and/or boolean functions of nodes. Bornholdt and Sneppen in \cite{bornholdt1998neutral} introduced an algorithm for RBN's evolution under the constraint of continuity. In the algorithm, a degree of a randomly chosen node is modified only when a network with this modification reaches the same attractor as the mother system did. Paczuski, Bassler and Corral in \cite{paczuski2000self} consider nodes of RBN as competing agents in a market. The evolution of network is maintained by a random change of the poorest player's strategy at each epoch of simulation. ADRR as a mechanism for transition of RBN to a critical state was proposed in \cite{bornholdt2000topological} and then extended in \cite{mlb} for annealed and quenched versions of a model. Both in  \cite{bornholdt2000topological} and \cite{mlb} ADRR changes a single randomly chosen node per iteration according to its average state on the attractor; the difference is that in \cite{mlb} boolean functions of nodes can also be changed at the end of the epoch.

	Simulation of a single epoch for adaptive RBN consists of two steps, search for an attractor and modification of a network according to chosen rules. Depending on the goal of simulation, it could be necessary to find a single attractor for a given initial state or all the attractors of a system. Zhang et. al. in \cite{zhang2007algorithms} propose algorithms to find singleton attractors of RBN using gene ordering. To find a single attractor of an arbitrary length, there are two main algorithms: Knuth's algorithm \cite{knuth1981art} (the example of its usage for RBN's simulation can be found in \cite{bhattacharjya1996median}) and the algorithm described in \cite{mlb} (we denote it as Liu-Bassler). We include the brief description of these algorithms in Section 3 of this paper as they serve as a basis for our parallel algorithm. The second category includes algorithms to find all attractors for a given network. These algorithms usually reformulate basic task as binary decision diagrams, satisfiability (SAT) or aggregation problems. For the asynchronous update mode, there was proposed an approach based on an exhaustive enumeration of subspaces of a network space implemented in Inet software package \cite{berntenis2013detection}. Some of the algorithms of this type allow parallel implementation (e.g. \cite{guo2014parallel}). Parallelization is performed by calculating different attractors (or sets of attractors) for the same network on different processors, and is not appropriate for ADRR having data dependency between epochs.

@@

	Due to the exponential slowing down by long attractor periods for initial connectivity $K_{ini} \geq 2$, usually it is necessary to put the restrictions on a maximum number of iterations to be examined during a single search for attractor (it is denoted as $T_{max}$). In \cite{bornholdt2000topological}, authors use $T_{max} = 1~000$ which allows them to study networks up to $1~024$ nodes. In \cite{mlb}, $T_{max}$ is set up to $100~000$ , and the maximum size of a studied network is equal to 400 nodes. Increase of a size of a network with fixed $T_{max}$ leads to skipping a significant number of the attractors. However, authors of \cite{bornholdt2000topological} and \cite{mlb} do not consider a percentage of attractors found, and how the results of simulation are influenced by skipped attractors.

@@

	Modeling the evolution of ARBN larger than several dozens of nodes is hampered by a significant growth of transient and attractor lengths. This growth extremely prolongs the time of a repeated calculation of nodes' states during search for an attractor (which is the most time consuming part of a modeling). To speed up modeling for large ARBNs, we propose to combine algorithms and techniques described in this Section. We use existing sequential algorithms for attractor's search with a limitation on a maximum number of iterations as a basis of the parallel algorithm, and extend ADRR to support larger number of nodes to be rewired.

@@

The modified algorithm of ARBN evolution consists of a set of consecutive epochs. In each epoch the network attractor is found and $N_{rew}$ nodes are rewired according to the modified ADRR. The maximum length of attractors $\mu_{max}$ is limited ether by predefined checkpoints in Liu-Bassler algorithm, or artificially in Knuth's algorithm in order to speed up the computations as described in Section \ref{sect:algorithm}. Due to this limitation only a fraction $R$ of the attractors is found. During the evolution process the system reaches a dynamical equilibrium, which can be observed by measuring network's mean in-degree connectivity $K$, defined in Eq. (\ref{eq:connectivity}). As the network evolves, $K$ tends to some steady-state value $K_{ss}$.

@@

	Intuitively, Liu-Bassler algorithm of attractor search would have shown better performance than the Knuth's one, as the latter requires at least $3 \mu$ state updates to find an attractor of length $\mu$. But in order to use Liu-Bassler algorithm one needs to choose the checkpoints correctly, which requires an a-priori knowledge of attractor and transient period distributions. This distribution depends on a network size, connectivity and structure (e. g. modular networks tend to have shorter attractors than non-modular ones \cite{Gorski2016}). We obtained an empirical distribution of attractor and transient lengths for different kinds of networks in our preliminary experiments with Knuth's algorithm. The checkpoints $T=[100, 200, 1~000, 2~000, ... \mu_{max}, 2 \mu_{max}]$ seem to be the best trade-off between the number of state updates and the percentage of attractors found. However, with these checkpoints the difference in the number of state updates between the two algorithms is negligible. We used Knuth's algorithm in our future experiments, as it is more general and requires no a-priori knowledge of the network's behaviour.

	We implemented the algorithm for parallel ARBN simulation, which supports both Liu-Bassler and Knuth's algorithms of attractor search. We used C++ programming language with CUDA extension. The implementation is cross-platform and requires a GPU with CUDA compute capability at least 1.2. We chose compressed sparse-row (CSR) matrix storage format to store matrix $A$ from Eq. (\ref{eq:spmv}) and implemented an sparse matrix-vector multiplication kernel as described in \cite{bell2008efficient}, as CSR kernel shows better performance on unstructured matrices with low number of non-zero elements than others do.

@@

	\paragraph{Speedup obtained by GPU acceleration.} We measured the execution times of both sequential and parallel versions of state updates. We used Liu-Bassler algorithm with checkpoints $T$ = $\{100,$ $200,$ $1~000,$ $2~000,$ $...,$ $10^7,$ $2 \cdot 10^7\}$, $\mu_{max} = 10^7$. To decrease the number of random samples required to compare the versions, we did not stop the epoch as soon as an attractor was found. Instead, we performed all $2 \cdot 10^7$ state updates per epoch. Such simulation corresponds to the worst-case scenario when no attractors were found, i. e. $R = 0\%$. The results presented in Figure \ref{fig:speedup} indicate that the GPU implementation is more efficient than the serial one for networks consisting of more than $200$ nodes.

	\begin{figure}[ht!]
		\begin{minipage}[t]{0.45\textwidth}
			\includegraphics[width=1.0\textwidth]{plots/speedup}
			\caption{Execution time of the GPU algorithm compared to the sequential one}
			\label{fig:speedup}
		\end{minipage}\hfill
		\begin{minipage}[t]{0.45\textwidth}
			\includegraphics[width=1.0\textwidth]{plots/n_rew}
			\caption{Evolution of mean connectivity for different $N_{rew}$; $N=1~000$}
			\label{fig:nrew}
		\end{minipage}
	\end{figure}

	\paragraph{Rewiring more than one node per epoch.} In order to determine how $N_{rew}$ influences the rate of convergence to a dynamical equilibrium, we performed $1~000$ epochs of evolution with $N=1~000$ and $N_{rew}=[1, 5]$. The results of these simulations are shown in Figure \ref{fig:nrew}. With increased value of $N_{rew}$ mean connectivity grows faster and the system reaches steady state at approximately 500th epoch for $N_{rew}=5$. Therefore, we conclude that by rewiring $N_{rew} > 1$ nodes per epoch one can lower the number of epochs required to reach steady state. We limited the attractor length to $\mu_{max}=10^7$ for these experiments.

	\paragraph{Limiting the maximum attractor length.} Introducing a limit on attractor length leads to a situation when an attractor can not be found. In these cases $N_{rew}$ nodes are rewired according to their activity during the last $\mu_{max}$ state updates. These states can belong to either a transient period or an attractor. However, the results of our experiments, presented in Tables \ref{table:percentage_of_attractors}, \ref{table:kss}, indicate that low percentage of attractors found does not influence the steady-state connectivity obtained as a result of the evolution process.

@@

	Using the approaches discussed above we were able perform simulations on networks consisting of up to $1~000$ nodes. Figure \ref{fig:evolution} shows the evolution of mean in-degree connectivity. The steady-state connectivity $K_{ss}$ decreases with growth of $N$, which corresponds to the results obtained in \cite{mlb}. The empirical distribution of attractor lengths for networks in steady state is shown in Figure \ref{fig:attractors}. Attractors seem to have power-law distribution as described in \cite{mlb}. The peaks near $\mu=10^7$  are caused by the limit on the attractor length $\mu_{max} = 10^7$.

	\begin{figure}[h]
		\begin{minipage}[t]{0.45\textwidth}
			\includegraphics[width=1.0\textwidth]{plots/evolution}
			\caption{Evolution of mean network connectivity for different $N$}
			\label{fig:evolution}
		\end{minipage}\hfill
		\begin{minipage}[t]{0.45\textwidth}
			\centering
			\includegraphics[width=1.0\textwidth]{plots/attractors}
			\caption{Empirical distribution of steady-state attractor lengths for different $N$}
			\label{fig:attractors}
		\end{minipage}
	\end{figure}
	%------------------------------------------------------------------------------
	\section{Conclusion}
	The enormous size of RBN ensemble makes them appropriate to model various complex systems but also raises significant challenges for their numerical investigation. Even a problem of search for a single attractor of RBN is computationally hard for a networks larger than several dozens of nodes because of an exponential growth of a number of iterations with an increase of a network size.

@@

	Using the algorithm, we performed simulation of networks up to $1~000$ nodes on a hybrid cluster with NVidia graphic cards. The parallel algorithm demonstrates increasing performance with a growth of a network size (for 1000-nodes network it is approximately 8 times faster than a sequential one). The results show that proposed modification of the rewiring rule allows to reduce a number of epochs to reach a steady state making calculations less time-consumable. We also present the results for a percentage of attractors found for different combinations of sizes of networks and maximum number of iterations examined during search for a single attractor. These values together with the results on mean steady-state connectivity can be used for a choice of values of checkpoints in the sequential algorithms of attractor's search.
